---
title: "Confounding"
output: html_notebook
---

#Confounding

Loading data:
```{r}
load(file = "rda/teams.rda")
```

```{r}
dat <- Teams %>%
  filter(yearID %in% 1961:2001)%>%
  mutate(R_game = R/G,
         BB_game = BB/G,
         HR_strata = (round(HR/G, 1)))%>%
  filter(HR_strata >= 0.4 & HR_strata <= 1.2)
dat %>%
  ggplot(aes(BB_game, R_game))+geom_point(alpha=0.5)+
  geom_smooth(method = "lm", se = F)+
  facet_wrap(~HR_strata)
```
```{r}
dat%>%
  group_by(HR_strata)%>%
  summarise(slope = cor(BB_game, R_game)*sd(R_game)/sd(BB_game))
```


RSS - Residual Sum Squares
```{r}
library(HistData)
rss <- function(beta0, beta1, data){
  resid <- galton_heights$son - (beta0+beta1*galton_heights$father)
  return(sum(resid^2))
}
beta1 = seq(0, 1, len=nrow(galton_heights))
results <- data.frame(beta1 = beta1, rss = sapply(beta1, rss, beta0 = 36))
results %>% ggplot(aes(beta1, rss)) + geom_line() + 
  geom_line(aes(beta1, rss), col=2)
```

#Linear Models

lm function

```{r}
dat <- Teams %>%
  filter(yearID %in% 1961:2001)%>%
  mutate(R_game = R/G,
         BB_game = BB/G,
         HR_game = ((HR/G)))
lm <- lm(R_game ~ BB_game + HR_game, data = dat)
summary(lm)
```


Least Square Errors are random variables:


```{r}
B <- 1000
N <- 100
lse <- replicate(B, {
  sample_n(galton_heights, N, replace = TRUE) %>% 
    lm(son ~ father, data = .) %>% .$coef 
})

lse <- data.frame(beta_0 = lse[1,], beta_1 = lse[2,]) 
lse %>% summarise(cor(beta_0, beta_1))
```

#More on linear models and tidyverse

##Tibbles

1. Tibbles display better:
```{r}
Teams
```
```{r}
as_tibble(Teams)
```

2. Subsets of data frame don't return a data frame, but the object class:
```{r}
class(Teams[,20])
```

However when we use tibbles, the result is another tibble:
```{r}
class(as_tibble(Teams[,20]))
```


3. Tibbles can have more complex objects:

For example we can put functions in a tibble:
```{r}
tibble(id = c(1,2,3), funcs = c(mean, median, sd))
```


4. We can group tibbles and the tidyverse functions are aware of them

However the lm function does not know how to deal with grouped data. To solve this we will need to use the function do:

##do function



```{r}
dat %>%
  group_by(HR)%>%
  do(fit = lm(R~BB, data = .))
```

```{r}
get_slope <- function(dato){
  fit <- lm(R ~ BB, data = dato)
  data.frame(slope = fit$coefficients[2],
             se = summary(fit)$coef[2,2])
}
dat %>%
  group_by(HR_strata)%>%
  do(get_slope(.))

```

##Broom Package! Designed to facilitate the use of models such as linear models with the tidyverse!!!

```{r}
dat%>%
  group_by(HR_strata)%>%
  do(tidy(lm(R~BB, data =.), conf.int = T))


dat%>%
  group_by(HR_strata)%>%
  do(tidy(lm(R_game~BB_game, data =.), conf.int = T))%>%
  filter(term == "BB_game")%>%
  select(HR_strata, estimate, conf.low, conf.high)%>%
  ggplot(aes(HR_strata, estimate, ymin = conf.low, ymax = conf.high))+
  geom_errorbar()+geom_point()+geom_hline(yintercept = 0.37, col= "red")

```



