---
title: "Correlation Coefficient"
output:
  html_document:
    df_print: paged
---
  
$\rho = \frac{1}{n}\sum\limits_{i=1}^{n}(\frac{x_i-\mu_x}{\sigma_x})*(\frac{y_i-\mu_y}{\sigma_y})$

```{r}
library(tidyverse)
library(HistData)
data("GaltonFamilies")
galton_heights <- GaltonFamilies%>%
  filter(childNum == 1 & gender == "male")%>%
  select(father, childHeight) %>%
  rename(son = childHeight)
galton_heights %>% summarise(cor(father, son))

```

Sample correlation is a random variable. And is the best predictor for the population correlation,

```{r}
set.seed(0)
R <- sample_n(galton_heights, 25, replace = T) %>%
  summarise(cor(father, son))
R
B <- 1000
R <- replicate(B, {
  sample_n(galton_heights, 25, replace = T) %>%
  summarise(r = cor(father, son))%>%
    .$r
})
data.frame(R)%>%
  ggplot(aes(R))+geom_histogram(binwidth = 0.05, color = "black")
```

The correlation can also be applied with the Central Limit Theorem, therefore the correlation coefficient can be model by
a normal variable with parameters:

$R \sim N(\rho,\sqrt(\frac{1-r^2}{N-2})$

Correlation is not always a good solution to infer trend between variables
 - One example is the Ascombe quartet, a conjunct of artificial data with different trends, in which all of them produce the same correlation coefficient
 
##Stratification

One way to predict son height based on the sample information and the father's height is to do a stratification of the population and compute the conditional average:

```{r}
conditional_average <- galton_heights %>%
  filter(round(father)==72)%>%
  summarise(avg = mean(son))%>%
  .$avg
conditional_average
```
This is larger than the average son.

stratification by boxplot can help to identify these stratas:

```{r}
galton_heights%>%
  mutate(father_strata = factor(round(father)))%>%
  ggplot(aes(x = father_strata,y = son))+
  geom_boxplot()+
  geom_point()
```

The means of each group seem to have a linear relationship:
```{r}
galton_heights%>%
  mutate(father_strata = factor(round(father)))%>%
  select(father_strata, son)%>%
  group_by(father_strata)%>%
  summarise(mean = mean(son))%>%
  mutate(father_strata = as.numeric(father_strata))%>%
  ggplot(aes(scale(father_strata), scale(mean)))+
  geom_point()+geom_abline(intercept = 0, slope = 0.5)
```

##Bivariate Normal Distribution

When a pair or variables (x,y) appear in a scatterplot with oval shape, it can be modelled as a bivariate normal distriution. Two variables highly correlated have a thin shape, or a circle if no correlation exists.
both x and y must be approximately normally distributed.

Exemple:

How much of the variance of y can be explained by x:
Ans: rho^2*100 [%]
```{r}
sd <- galton_heights %>%
  summarise(sd = sd(son))%>%
  pull(sd)
1-(1-0.5^2)*sd^2/sd^2 

```

